Accelerated Inference for Partially Observed Markov Processes using Automatic Differentiation

### 2024-06-14 paper/appendix/convergence.tex
update from overleaf
another day, another dollar

\begin{lem}[Concentration of Measure for Gradient Estimate]
    \label{lemma:grad_bound}
    Consider the gradient estimate obtained by MOP-1, which we know by Theorem \ref{thm:mop-grad-consistency} is consistent for the score, where $\theta = \phi$. For $||\nabla_\theta \hat{\loglik}^1(\theta) - \nabla_\theta \loglik(\theta)||_2$ to be bounded by $\epsilon$ with probability $1-\delta$, we require
    \begin{align}
    J > \max\left\{2G(\theta)\frac{r_N\sqrt{p}}{\epsilon}\left(1+h^{-1}\left(\log\left(\frac{2p}{\delta}\right)\right)\right), 8G(\theta)^2\beta_N^2\frac{p\log(2p/\delta)}{\epsilon^2}\right\},
    \end{align}
    where $NG'(\theta) \leq G(\theta)$ are defined in Assumptions \ref{assump:bounded-measurement} and \ref{assump:local-bounded-derivative}, $h(t) = \frac{1}{2}(t - \log(1+t))$, and $\beta_N$ and $r_N$ are two additional finite model-specific constants that do not depend on $J$, but do depend on $N$ and $p$, as defined in \cite{delMoral11}. 
Equivalently, with probability at least $1-\delta$, it holds that
    \begin{align}
        ||\nabla_\theta \hat\ell^1(\theta) - \nabla_\theta \ell(\theta)||_2 \leq G(\theta)\left(\frac{r_N\sqrt{p}}{J}(1+h^{-1}(\log(2p/\delta))) + \sqrt{\frac{2p\log(2p/\delta)}{J}}\beta_N\right).
    \end{align}
\end{lem}

@@

Now we apply the Hoeffding-type concentration inequality from Del Moral and Rio \cite{delMoral11} and a union bound over each $\varphi_n^i(x_{n,j}^{F,\theta})$, totaling $N$ timesteps and $p$ parameters, to find that
\begin{align}
    \max_{n=1,...,N} \left\lVert\frac{1}{J}\sum_{j=1}^J\nabla_\theta \log g_{n,k_j}^{A,\theta} - \nabla_\theta \ell_n(\theta) \right\rVert_{\infty} \leq G'(\theta)\left(\frac{r_N}{J}(1+h^{-1}(t)) + \sqrt{\frac{2t}{J}}\beta_N \right)
\end{align}
with probability at least $1-2Np\exp(-t)$. Although the above concentration inequality only considers the error from the expectation under the filtering distribution, we invoke the consistency of MOP-$1$ shown in Theorem \ref{thm:mop-grad-consistency} to establish that the expectation under the filtering distribution is in fact the score. It therefore holds that with the same probability, that when summing over $N$, as $NG'(\theta) \leq G(\theta)$, 
\begin{align}
    \left\lVert\frac{1}{J}\sum_{j=1}^J\sum_{n=1}^N\nabla_\theta \log g_{n,k_j}^{A,\theta} - \nabla_\theta \ell(\theta) \right\rVert_{\infty} 
    &\leq G'(\theta)N\left(\frac{r_N}{J}(1+h^{-1}(t)) + \sqrt{\frac{2t}{J}}\beta_N \right)\\
    &\leq G(\theta)\left(\frac{r_N}{J}(1+h^{-1}(t)) + \sqrt{\frac{2t}{J}}\beta_N \right).
\end{align}
We split the $\delta$ failure probability among these $2Np$ terms, to find $\delta\leq2Np\exp(-t)$, and therefore, $t\leq\log(2Np/\delta)$, where $h(t) = \frac{1}{2}(t - \log(1+t))$. 
The two additional model-specific parameters are $\beta_t$ and $r_t$, which do not depend on $J$. 
The analogous bound for the 2-norm follows from scaling the right-hand side by $\sqrt{p}$, to require 
\begin{align}
    ||\nabla_\theta \hat\ell(\theta) - \nabla_\theta \ell(\theta)||_2 \leq G(\theta)\left(\frac{r_N\sqrt{p}}{J}(1+h^{-1}(\log(2p/\delta))) + \sqrt{\frac{2p\log(2p/\delta)}{J}}\beta_N\right).
\end{align}
We therefore need 
\begin{align}
    J > \max\left\{2G(\theta)\frac{r_N\sqrt{p}}{\epsilon}\left(1+h^{-1}\left(\log\left(\frac{2p}{\delta}\right)\right)\right), 8G(\theta)^2\beta_N^2\frac{p\log(2p/\delta)}{\epsilon^2}\right\}.
\end{align}

@@

It is known that the minimum eigenvalue of a symmetric matrix is concave. Therefore, it suffices to show that the first inequality in the below expression
\begin{equation}
    0 < \sum_{j=1}^J \lambda_{\min} (H_j) \leq  \lambda_{\min}\left(\sum_{j=1}^J H_j\right) = \lambda_{\min} (\hat{H})
\end{equation}
holds with high probability.
We apply the particle Hoeffding concentration inequality from \cite{delMoral11} to find that  
\begin{align}
    \frac{1}{J}\sum_{j=1}^J \lambda_{\min}(H_j) - \E_{\tilde{\pi}_t}\lambda_{\min}(H_j) &= \frac{1}{J}\sum_{j=1}^J \lambda_{\min}(H_j) - \gamma' \geq -\frac{r_t}{J}(1+h^{-1}(t)) - \sqrt{\frac{2t}{J}}\beta_t \\
    \sum_{j=1}^J \lambda_{\min}(H_j) &\geq -r_t(1+h^{-1}(t)) - \sqrt{2tJ}\beta_t + J\gamma',
\end{align}
with probability at least $1-\exp(-t)$. Here, $h(t) = \frac{1}{2}(t - \log(1+t))$. 
The two additional model-specific parameters are $\beta_t$ and $r_t$, which do not depend on $J$.

We additionally require, for $c \in (0, \sum_{j=1}^J \E\lambda_{\min}(H_j))$,
\begin{align}
    \sum_{j=1}^J \lambda_{\min}(H_j) \geq -r_t(1+h^{-1}(t)) - \sqrt{2tJ}\beta_t + J\gamma' \geq c, \\
    J\gamma' \geq c + r_t(1+h^{-1}(t)) + \sqrt{2tJ}\beta_t.
\end{align}
We therefore require \ed{WE SAY ``REQUIRE'' AT VARIOUS POINTS, BUT THAT SUGGESTS NECESSITY WHEREAS I SUPPOSE WE MEAN SUFFICIENCY. I EDITED ONE OR TWO OF THESE.}
\begin{equation}
J > \max\left\{\frac{2r_t(1+h^{-1}(t)) + 2c}{\gamma'}, \frac{2(2t\beta_t^2+c)^2}{\gamma'^2}\right\} \geq  \frac{r_t(1+h^{-1}(t))}{\gamma'} + \sqrt{2tJ}\beta_t/\gamma' + c/\gamma'    
\end{equation}
for $\hat{H}(\theta)$ to be invertible and positive definite with minimum eigenvalue greater than or equal to $c$ with probability at least $1-\exp(-t)$.

@@

    f(\theta_m+\eta p_m) \leq f(\theta_m) + \eta\beta p_m^Tg(\theta_m),

@@

By an argument found in the beginning of the proof of Theorem 6 in \cite{mahoney16}, we have that choosing $J$ such that $||\nabla_\theta\hat{\ell}(\theta_m) - \nabla_\theta \ell(\theta_m)|| \leq \epsilon$ and $\lambda_{\min}(H(\theta_m)) \geq c>0$ for each $m$, yields

@@

    f(\theta_\eta)-f(\theta_m) \leq \eta p_m^Tg(\theta_m) + \epsilon\eta||p_m|| + \eta^2 \Gamma ||p_m||^2 / 2,

@@

    p_m^Tg(\theta_m) = -p_m^TH(\theta_m)p_m \geq -c||p_m||^2,

@@

    f(\theta_\eta)-f(\theta_m) \leq -\eta p_m^TH(\theta_m)p_m + \epsilon\eta||p_m|| + \eta^2 \Gamma ||p_m||^2 / 2,

@@

    -\eta p_m^TH(\theta_m)p_m + \epsilon\eta||p_m|| + \eta^2 \Gamma ||p_m||^2 / 2 &\leq \eta \beta p_m^Tg(\theta_m) = - \eta \beta p_m^TH(\theta_m)p_m \\
    \epsilon||p_m|| + \eta \Gamma ||p_m||^2 / 2 &\leq (1- \beta) p_m^TH(\theta_m)p_m \\
    \epsilon + \eta \Gamma ||p_m|| / 2 &\leq c(1- \beta) ||p_m||.

@@

    \eta \leq \frac{c(1-\beta)}{\Gamma}, \;\; \epsilon \leq \frac{c(1-\beta)}{2\Gamma}||g(\theta_m)|| \leq \frac{c(1-\beta)}{2}||p_m||,

@@

||g(\theta_m)|| - ||\nabla_\theta f(\theta_m)|| \leq ||g(\theta_m) - \nabla_\theta f(\theta_m)|| \leq \epsilon => ||\nabla_\theta f(\theta_m)|| \geq ||g(\theta_m)|| - \epsilon

@@

||\nabla_\theta f(\theta_m)||-||g(\theta_m)|| \leq ||\nabla_\theta f(\theta_m)-g(\theta_m)|| \leq \epsilon => ||g(\theta_m)|| \geq ||\nabla_\theta f(\theta_m)|| - \epsilon.

@@

If the algorithm terminates and $||g(\theta_m)|| \leq \sigma \epsilon$, we can derive 

@@

    ||\nabla_\theta f(\theta_m)|| \leq ||g(\theta_m)|| + \epsilon = \sigma\epsilon+\epsilon = (\sigma+1)\epsilon.

@@

If the algorithm does not terminate, then $||g(\theta_m)|| > \sigma \epsilon$. 

@@

    \epsilon \geq ||g(\theta_m) - \nabla_\theta f(\theta_m)|| &\geq& ||g(\theta_m)|| - ||\nabla_\theta f(\theta_m)|| 

@@

    ||\nabla_\theta f(\theta_m)|| + \epsilon &\geq& ||g(\theta_m)|| \geq \sigma \epsilon 

@@

    ||\nabla_\theta f(\theta_m)|| &\geq& \sigma \epsilon - \epsilon = (\sigma - 1)\epsilon 

@@

    \frac{||\nabla_\theta f(\theta_m)||}{\sigma-1} &\geq& \epsilon,

@@

    ||\nabla_\theta f(\theta_m)|| - \epsilon
    &\geq&  ||\nabla_\theta f(\theta_m)|| - \frac{||\nabla_\theta f(\theta_m)||}{\sigma-1} 

@@

    &=& \left(1-\frac{1}{\sigma-1}\right)||\nabla_\theta f(\theta_m|| 

@@

    &=& \frac{\sigma-2}{\sigma-1}||\nabla_\theta f(\theta_m)|| 

@@

    &\geq& \frac{2}{3}||\nabla_\theta f(\theta_m)||.

@@

Since $||A^{-1}|| = 1/\sigma_{\min}(A)$,

@@

    p_m^TH(\theta_m)p_m &= (-(H(\theta_m))^{-1}g(\theta_m))^TH(\theta_m)(-(H(\theta_m))^{-1}g(\theta_m)) \\

@@

    &\geq \frac{1}{c}||g(\theta_m)||^2 \\
    &\geq \frac{1}{c}(||\nabla_\theta f(\theta_m)|| - \epsilon)^2 \\
    &\geq \frac{4}{9c}||\nabla_\theta f(\theta_m)||^2.

@@

    f(\theta_m) - f(\theta^*) \leq \frac{1}{2\gamma}||\nabla_\theta f(\theta_m)||^2,

@@

    f(\theta_m) - f(\theta^*) \leq \frac{1}{2\gamma}||\nabla_\theta f(\theta_m)||^2 \leq \frac{9c}{4}\frac{1}{2\gamma}p_m^TH(\theta_m)p_m,

@@

    \frac{8\gamma}{9c}(f(\theta_m) - f(\theta^*)) \leq \frac{4}{9c}||\nabla_\theta f(\theta_m)||^2 \leq p_m^TH(\theta_m)p_m,

@@

    f(\theta_m) - f(\theta^*) \leq \frac{9c}{8\gamma}p_m^TH(\theta_m)p_m,

@@

    -\frac{8\gamma}{9c}(f(\theta_m) - f(\theta^*)) \geq -\frac{4}{9c}||\nabla_\theta f(\theta_m)||^2 \geq -p_m^TH(\theta_m)p_m.

@@

    f(\theta_{m+1})-f(\theta_m) &\leq -\eta p_m^TH(\theta_m)p_m + \epsilon\eta||p_m|| + \eta^2 \Gamma ||p_m||^2 / 2 \\
    &\leq -\eta\beta p_m^TH(\theta_m)p_m \\
    &\leq -\eta\beta\frac{8\gamma}{9c}(f(\theta_m) - f(\theta^*)).

@@

    &\leq f(\theta_m)- f(\theta^*) -\eta\beta\frac{8\gamma}{9c}(f(\theta_m) - f(\theta^*)) \\
    &= (1-\eta\beta\frac{8\gamma}{9c})(f(\theta_m) - f(\theta^*)).

>>>>

\begin{lem}[Concentration of Measure for Gradient Estimate]
    \label{lemma:grad_bound}
    Consider the gradient estimate obtained by MOP-1, which we know by Theorem \ref{thm:mop-grad-consistency} is consistent for the score, where $\theta = \phi$. For $\|\nabla_\theta \hat{\loglik}^1(\theta) - \nabla_\theta \loglik(\theta)\|_2$ to be bounded by $\epsilon$ with probability $1-\delta$, we require
    \begin{align}
    J > \max\left\{2G(\theta)\frac{r_N\sqrt{p}}{\epsilon}\left(1+h^{-1}\left(\log\left(\frac{2p}{\delta}\right)\right)\right), 8G(\theta)^2\beta_N^2\frac{p\log(2p/\delta)}{\epsilon^2}\right\},
    \end{align}
    where $NG'(\theta) \leq G(\theta)$ are defined in Assumptions \ref{assump:bounded-measurement} and \ref{assump:local-bounded-derivative}, $h(t) = \frac{1}{2}(t - \log(1+t))$, and $\beta_N$ and $r_N$ are two additional finite model-specific constants that do not depend on $J$, but do depend on $N$ and $p$, as defined in \cite{delMoral11}. 
Equivalently, with probability at least $1-\delta$, it holds that
    \begin{align}
        \|\nabla_\theta \hat\ell^1(\theta) - \nabla_\theta \ell(\theta)\|_2 \leq G(\theta)\left(\frac{r_N\sqrt{p}}{J}(1+h^{-1}(\log(2p/\delta))) + \sqrt{\frac{2p\log(2p/\delta)}{J}}\beta_N\right).
    \end{align}
\end{lem}

@@

Now we apply the Hoeffding-type concentration inequality from Del Moral and Rio \cite{delMoral11} and a union bound over each $\varphi_n^i(x_{n,j}^{F,\theta})$, totaling $N$ timesteps and $p$ parameters, to find that
\begin{align}
    \max_{n=1,...,N} \left\lVert\frac{1}{J}\sum_{j=1}^J\nabla_\theta \log g_{n,k_j}^{A,\theta} - \nabla_\theta \ell_n(\theta) \right\rVert_{\infty} \leq G'(\theta)\left(\frac{r_N}{J}(1+h^{-1}(t)) + \sqrt{\frac{2t}{J}}\beta_N \right)
\end{align}
with probability at least $1-2Np\exp(-t)$. Although the above concentration inequality only considers the error from the expectation under the filtering distribution, we invoke the consistency of MOP-$1$ shown in Theorem \ref{thm:mop-grad-consistency} to establish that the expectation under the filtering distribution is in fact the score. It therefore holds that with the same probability, that when summing over $N$, as $NG'(\theta) \leq G(\theta)$, 
\begin{align}
    \left\lVert\frac{1}{J}\sum_{j=1}^J\sum_{n=1}^N\nabla_\theta \log g_{n,k_j}^{A,\theta} - \nabla_\theta \ell(\theta) \right\rVert_{\infty} 
    &\leq G'(\theta)N\left(\frac{r_N}{J}(1+h^{-1}(t)) + \sqrt{\frac{2t}{J}}\beta_N \right)\\
    &\leq G(\theta)\left(\frac{r_N}{J}(1+h^{-1}(t)) + \sqrt{\frac{2t}{J}}\beta_N \right).
\end{align}
We split the $\delta$ failure probability among these $2Np$ terms, to find $\delta\leq2Np\exp(-t)$, and therefore, $t\leq\log(2Np/\delta)$, where $h(t) = \frac{1}{2}(t - \log(1+t))$. 
The two additional model-specific parameters are $\beta_t$ and $r_t$, which do not depend on $J$. 
The analogous bound for the 2-norm follows from scaling the right-hand side by $\sqrt{p}$, to require 
\begin{align}
    \|\nabla_\theta \hat\ell(\theta) - \nabla_\theta \ell(\theta)\|_2 \leq G(\theta)\left(\frac{r_N\sqrt{p}}{J}(1+h^{-1}(\log(2p/\delta))) + \sqrt{\frac{2p\log(2p/\delta)}{J}}\beta_N\right).
\end{align}
We therefore need 
\begin{align}
    J > \max\left\{2G(\theta)\frac{r_N\sqrt{p}}{\epsilon}\left(1+h^{-1}\left(\log\left(\frac{2p}{\delta}\right)\right)\right), 8G(\theta)^2\beta_N^2\frac{p\log(2p/\delta)}{\epsilon^2}\right\}.
\end{align}

@@

It is known that the minimum eigenvalue of a symmetric matrix is concave. Therefore, it suffices to show that the first inequality in the below expression
\begin{equation}
    0 < \sum_{j=1}^J \lambda_{\min} (H_j) \leq  \lambda_{\min}\left(\sum_{j=1}^J H_j\right) = \lambda_{\min} (\hat{H})
\end{equation}
holds with high probability.
We apply the particle Hoeffding concentration inequality from \cite{delMoral11} to find that  
\begin{align}
    \frac{1}{J}\sum_{j=1}^J \lambda_{\min}(H_j) - \E_{\tilde{\pi}_t}\lambda_{\min}(H_j) &= \frac{1}{J}\sum_{j=1}^J \lambda_{\min}(H_j) - \gamma' \geq -\frac{r_t}{J}\big(1+h^{-1}(t)\big) - \sqrt{\frac{2t}{J}}\beta_t \\
    \sum_{j=1}^J \lambda_{\min}(H_j) &\geq -r_t\big(1+h^{-1}(t)\big) - \sqrt{2tJ}\beta_t + J\gamma',
\end{align}
with probability at least $1-\exp(-t)$. Here, $h(t) = \frac{1}{2}(t - \log(1+t))$. 
The two additional model-specific parameters are $\beta_t$ and $r_t$, which do not depend on $J$.

We additionally require, for $c \in \big(0, \sum_{j=1}^J \E\lambda_{\min}(H_j)\big)$,
\begin{align}
    \sum_{j=1}^J \lambda_{\min}(H_j) \geq -r_t\big(1+h^{-1}(t)\big) - \sqrt{2tJ}\beta_t + J\gamma' \geq c, \\
    J\gamma' \geq c + r_t\big(1+h^{-1}(t)\big) + \sqrt{2tJ}\beta_t.
\end{align}
It is therefore sufficient to have
\begin{equation}
J > \max\left\{\frac{2r_t\big(1+h^{-1}(t)\big) + 2c}{\gamma'}, \frac{2(2t\beta_t^2+c)^2}{\gamma'^2}\right\} \geq  \frac{r_t\big(1+h^{-1}(t)\big)}{\gamma'} + \sqrt{2tJ}\beta_t/\gamma' + c/\gamma'    
\end{equation}
for $\hat{H}(\theta)$ to be invertible and positive definite with minimum eigenvalue greater than or equal to $c$ with probability at least $1-\exp(-t)$.

@@

    f(\theta_m+\eta p_m) \leq f(\theta_m) + \eta\beta \, p_m^T \, g(\theta_m),

@@

By an argument found in the beginning of the proof of Theorem 6 in \cite{mahoney16}, we have that choosing $J$ such that $\|\nabla_\theta\hat{\ell}(\theta_m) - \nabla_\theta \ell(\theta_m)\| \leq \epsilon$ and $\lambda_{\min}(H(\theta_m)) \geq c>0$ for each $m$, yields

@@

    f(\theta_\eta)-f(\theta_m) \leq \eta p_m^T\, g(\theta_m) + \epsilon\eta\|p_m\| + \eta^2 \Gamma \|p_m\|^2 / 2,

@@

    p_m^Tg(\theta_m) = -p_m^T\, H(\theta_m)\, p_m \geq -c\|p_m\|^2,

@@

    f(\theta_\eta)-f(\theta_m) \leq -\eta p_m^TH(\theta_m)\, p_m + \epsilon\eta\|p_m\| + \eta^2 \Gamma \|p_m\|^2 / 2,

@@

    -\eta p_m^TH(\theta_m)p_m + \epsilon\eta\|p_m\| + \eta^2 \Gamma \|p_m\|^2 / 2 &\leq \eta \beta p_m^Tg(\theta_m) = - \eta \beta p_m^TH(\theta_m)p_m \\
    \epsilon\|p_m\| + \eta \Gamma \|p_m\|^2 / 2 &\leq (1- \beta) p_m^TH(\theta_m)p_m \\
    \epsilon + \eta \Gamma \|p_m\| / 2 &\leq c(1- \beta) \|p_m\|.

@@

    \eta \leq \frac{c(1-\beta)}{\Gamma}, \;\; \epsilon \leq \frac{c(1-\beta)}{2\Gamma}\|g(\theta_m)\| \leq \frac{c(1-\beta)}{2}\|p_m\|,

@@

\|g(\theta_m)\| - \|\nabla_\theta f(\theta_m)\| \leq \|g(\theta_m) - \nabla_\theta f(\theta_m)\| \leq \epsilon \implies \|\nabla_\theta f(\theta_m)\| \geq \|g(\theta_m)\| - \epsilon

@@

\|\nabla_\theta f(\theta_m)\|-\|g(\theta_m)\| \leq \|\nabla_\theta f(\theta_m)-g(\theta_m)\| \leq \epsilon \implies \|g(\theta_m)\| \geq \|\nabla_\theta f(\theta_m)\| - \epsilon.

@@

If the algorithm terminates and $\|g(\theta_m)\| \leq \sigma \epsilon$, we can derive 

@@

    \|\nabla_\theta f(\theta_m)\| \leq \|g(\theta_m)\| + \epsilon = \sigma\epsilon+\epsilon = (\sigma+1)\epsilon.

@@

If the algorithm does not terminate, then $\|g(\theta_m)\| > \sigma \epsilon$. 

@@

    \epsilon \geq \|g(\theta_m) - \nabla_\theta f(\theta_m)\| &\geq& \|g(\theta_m)\| - \|\nabla_\theta f(\theta_m)\| 

@@

    \|\nabla_\theta f(\theta_m)\| + \epsilon &\geq& \|g(\theta_m)\| \geq \sigma \epsilon 

@@

    \|\nabla_\theta f(\theta_m)\| &\geq& \sigma \epsilon - \epsilon = (\sigma - 1)\epsilon 

@@

    \frac{\|\nabla_\theta f(\theta_m)\|}{\sigma-1} &\geq& \epsilon,

@@

    \|\nabla_\theta f(\theta_m)\| - \epsilon
    &\geq&  \|\nabla_\theta f(\theta_m)\| - \frac{\|\nabla_\theta f(\theta_m)\|}{\sigma-1} 

@@

    &=& \left(1-\frac{1}{\sigma-1}\right)\|\nabla_\theta f(\theta_m\| 

@@

    &=& \frac{\sigma-2}{\sigma-1}\|\nabla_\theta f(\theta_m)\| 

@@

    &\geq& \frac{2}{3}\|\nabla_\theta f(\theta_m)\|.

@@

Since $\|A^{-1}\| = 1/\sigma_{\min}(A)$,

@@

    p_m^TH(\theta_m)p_m &= \big(-(H(\theta_m))^{-1}g(\theta_m)\big)^TH(\theta_m)\big(-(H(\theta_m))^{-1}g(\theta_m)\big) \\

@@

    &\geq \frac{1}{c}\|g(\theta_m)\|^2 \\
    &\geq \frac{1}{c}\big(\|\nabla_\theta f(\theta_m)\| - \epsilon\big)^2 \\
    &\geq \frac{4}{9c}\|\nabla_\theta f(\theta_m)\|^2.

@@

    f(\theta_m) - f(\theta^*) \leq \frac{1}{2\gamma}\big\|\nabla_\theta f(\theta_m)\big\|^2,

@@

    f(\theta_m) - f(\theta^*) \leq \frac{1}{2\gamma}\big\|\nabla_\theta f(\theta_m)\big\|^2 \leq \frac{9c}{4}\frac{1}{2\gamma}\, p_m^TH(\theta_m)\, p_m,

@@

    \frac{8\gamma}{9c}\big(f(\theta_m) - f(\theta^*)\big) \leq \frac{4}{9c}\, \big\|\nabla_\theta f(\theta_m)\big\|^2 \leq p_m^TH(\theta_m)\, p_m,

@@

    f(\theta_m) - f(\theta^*) \leq \frac{9c}{8\gamma}\, p_m^TH(\theta_m)\, p_m,

@@

    -\frac{8\gamma}{9c}\big(f(\theta_m) - f(\theta^*)\big) \geq -\frac{4}{9c}\, \big\|\nabla_\theta f(\theta_m)\big\|^2 \geq -p_m^TH(\theta_m)\, p_m.

@@

    f(\theta_{m+1})-f(\theta_m) &\leq -\eta \, p_m^TH(\theta_m)p_m + \epsilon\eta\|p_m\| + \eta^2 \Gamma \|p_m\|^2 / 2 \\
    &\leq -\eta\beta \,  p_m^TH(\theta_m)\, p_m \\
    &\leq -\eta\beta \, \frac{8\gamma}{9c}\, \big(f(\theta_m) - f(\theta^*)\big).

@@

    &\leq f(\theta_m)- f(\theta^*) -\eta\beta\, \frac{8\gamma}{9c} \, \big(f(\theta_m) - f(\theta^*)\big) \\
    &= \Big(1-\eta\beta\frac{8\gamma}{9c}\Big)\big(f(\theta_m) - f(\theta^*)\big).
